---
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
<span style="color:#008B8B ;font-family:Times New Roman; font-size:14;">
<h2>Analyzing Credit Card Limits</h2>
<h2>BDA503: Default Of Credit Card Clients</h2>
<p>================================================================<br>
Group Name: <span style="color:black;font-size:12"> Paranormal Distribution </span><br>
Data Set: https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients<br>
================================================================</p>
</span>

### Load Required Modules
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#if you dont use these libraries before, please install them first!
#install.packages('ggplot2')
#install.packages('RColorBrewer')
#install.packages('gridExtra')
#install.packages('corrplot')
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(corrplot)
```
### Load and Investigate Data
```{r, warning=FALSE, comment=''}
ccdata <- read.csv("../input/UCI_Credit_Card.csv")
#Before tidying data set
dim(ccdata)
#Before tidying data set
str(ccdata)
#Before tidying data set
summary(ccdata)
```
### Reshaping Data 1: Creating New Variables
```{r, warning=FALSE, comment=''}
ccdata$workstate <- ""
ccdata$genderH <- ""
ccdata$educationH <- ""
ccdata$maritalH <- ""

for (i in 1:nrow(ccdata)) {
  if ((ccdata[i,7] + ccdata[i,8] +ccdata[i,9]+ccdata[i,10] +ccdata[i,11]+ccdata[i,12]) <= 0){
    ccdata[i,26] <- "YES"  
  }
  else {
    ccdata[i,26] <- "NO"         
  }
}

for (i in 1:nrow(ccdata)) {
  if (ccdata[i,3] == 1) {
    ccdata[i,27] <- "Male"  
  }
  else {
    ccdata[i,27] <- "Female"         
  }
}

for (i in 1:nrow(ccdata)) {
  if (ccdata[i,4] == 1) {
    ccdata[i,28] <- "Graduate"
  } else if (ccdata [i,4] == 2) {
    ccdata[i,28] <- "University" 
  } else if (ccdata [i,4] == 3) {
    ccdata[i,28] <- "High School" 
  } else {
    ccdata[i,28] <- "Unknown" 
  }
}

for (i in 1:nrow(ccdata)) {
  if(ccdata[i,5] == 1) {
    ccdata[i,29] <- "Married"
  } else if (ccdata[i,5] == 2) {
    ccdata[i,29] <- "Single"
  } else {
    ccdata[i,29] <- "Other"
  }
}
```
### Cut a Variable
```{r, comment=''}
ccdata$AGE.bucket<-cut(ccdata$AGE,c(10,20,30,40,50,60,70))
```
### Reshaping Data 2: Converting Quantitative Variables To New Factorial Variables
```{r, comment=''}
ccdata$workstate <-factor(ccdata$workstate)
ccdata$SEX <-factor(ccdata$SEX)
ccdata$EDUCATION <- factor(ccdata$EDUCATION)
ccdata$MARRIAGE <- factor(ccdata$MARRIAGE)
ccdata$AGEf <- factor(ccdata$AGE)
ccdata$default.payment.next.month<-factor(ccdata$default.payment.next.month)
ccdata$genderH <- factor(ccdata$genderH)
ccdata$educationH <- factor(ccdata$educationH)
ccdata$maritalH <- factor(ccdata$maritalH)
```
### Reshaping Data 3: Removing Variables After Reshaping 1
```{r, comment=''}
ccdata = subset(ccdata, select = -c(PAY_0,PAY_2,PAY_3,PAY_4,PAY_5,PAY_6))
```
### Investigate Data Again
```{r, comment=''}
#After tidying data set
dim(ccdata)
#After tidying data set
str(ccdata)
#After tidying data set
summary(ccdata)
```
### Balance Limit by Gender, Education, Work State
```{r, comment=''}
# Balance limits by gender and education
d1 <- ggplot(ccdata, aes(factor(genderH), (LIMIT_BAL/1000), fill=educationH)) + 
  geom_boxplot() +
  xlab("Gender") + 
  ylab("BLimit(x1000 NT$)") + 
  scale_fill_brewer(palette = "Accent")

# Balance limits by education and gender
d2 <- ggplot(ccdata, aes(factor(educationH), (LIMIT_BAL/1000), fill=genderH)) + 
  geom_boxplot() +
  xlab("Education") + 
  ylab("BLimit(x1000 NT$)") + 
  scale_fill_brewer(palette = "Paired")

# Balance limits by workstate and education
d3 <-ggplot(ccdata, aes(factor(educationH), (LIMIT_BAL/1000), fill=workstate)) + 
  geom_boxplot() +
  xlab("Education") + 
  ylab("BLimit(x1000 NT$)") 

grid.arrange(d1, d2, d3)

#Result: When we compare the balance limits with gender, education and work status, we saw that gender has no effects on balance limit determination process of bank while the education level is has a positive effect on this process. Also it can be seen that work status is a very important factor at balance limit determination.
```
### Relation Between Marital Status & Balance Limits By Gender
```{r, warning=FALSE, comment=''}
ggplot(ccdata, aes(factor(maritalH), (LIMIT_BAL/1000), fill=genderH)) + 
  geom_boxplot() +
  xlab("Marital Status") + 
  ylab("Balance Limit ( x 1000 NT$)") + 
  coord_cartesian(ylim = c(0,350)) +
  scale_fill_brewer(palette = "Paired")
#Result: By this graph, we can see it again that, there is no change at females side such as balance limits depending on their marital status, however it changes a lot of things on males’ side starting with the expenditures which is the reason on increased balance limits. 
```
### Histogram of Limit Balance & Default Payment
```{r, message=FALSE, warning=FALSE, comment=''}
ggplot(aes(x = ccdata$LIMIT_BAL/1000), data = ccdata) +
geom_histogram(aes(fill = ccdata$default.payment.next.month)) +
xlab("Balance Limit x 1000") +
ylab("Count") +
scale_fill_discrete(name="Default Payment Next Month",
                      breaks=c(0, 1),
                      labels=c("No", "Yes")) +
xlim(c(0,750)) +
facet_wrap(~educationH)

#Result: Balance limits and count of defaulted clients are almost same for University and Graduate Level. Additionally, the ratio of defaulted clients at high school level seems almost the same as the university and graduate levels.
```

### Relation Between Education & Default Payment
```{r, message=FALSE, warning=FALSE, comment=''}
d1 <- ggplot(ccdata, aes(x=default.payment.next.month)) + 
  geom_histogram(stat="count",color='red',fill='orange') +
  xlab("Default Payment Status") + ylab("Customer Count") + 
  facet_wrap(~educationH)

d2 <- ggplot(ccdata, aes(x=default.payment.next.month),aes(y=stat_count(gender))) + 
  geom_bar(aes(fill=factor(ccdata$educationH))) +
  xlab("Default Payment Status")+ylab("Customer Count") + 
  facet_wrap(~genderH)+
  scale_fill_discrete(name="Education")

grid.arrange(d1, d2, ncol=1)

#Result: There is no effect of education level on default event occurrence.
```
### Balance Limits By Age Groups & Education
```{r, comment=''}
ggplot(data = subset(ccdata,!is.na(AGE.bucket)), aes(factor(educationH), (LIMIT_BAL/1000), fill=AGE.bucket)) + 
  geom_boxplot() +
  xlab("Education") + 
  ylab("Balance Limit ( x 1000 NT$)") + 
  coord_cartesian(ylim = c(0,500)) +
  scale_fill_brewer(palette = "Accent")
#Result: Reflection of the education levels on determination of balance limits for clients is increasing by later ages when we compare the averages of each age buckets with each other grouped by their education levels.
```
### Data Heading Conversation of Customer Bills
```{r, comment=''}
ccdata$SEP2005<-ccdata$BILL_AMT1
ccdata$AUG2005<-ccdata$BILL_AMT2
ccdata$JUL2005<-ccdata$BILL_AMT3
ccdata$JUN2005<-ccdata$BILL_AMT4
ccdata$MAY2005<-ccdata$BILL_AMT5
ccdata$APR2005<-ccdata$BILL_AMT6
```
### Expdenditure By Months
```{r, comment=''}
apr <- ggplot(aes(x=AGE,y=APR2005/1000),data=ccdata) +
  xlab("Age") + 
  ylab("Amount of Bill in April") +
  coord_cartesian(xlim = c(21,60),ylim = c(0,700))+  
  geom_jitter(alpha=0.3, color="orange") + geom_smooth(stat='summary', fun.y=mean)

may <- ggplot(aes(x=AGE,y=MAY2005/1000),data=ccdata) +
  xlab("Age") + 
  ylab("Amount of Bill in May") +
  coord_cartesian(xlim = c(21,60),ylim = c(0,700))+  
  geom_jitter(alpha=0.3, color="blue") + geom_smooth(stat='summary', fun.y=mean)

jun <- ggplot(aes(x=AGE,y=JUN2005/1000),data=ccdata) +
  xlab("Age") + 
  ylab("Amount of Bill in June") +
  coord_cartesian(xlim = c(21,60),ylim = c(0,700))+  
  geom_jitter(alpha=0.3, color="green") + geom_smooth(stat='summary', fun.y=mean)

jul <- ggplot(aes(x=AGE,y=JUL2005/1000),data=ccdata) +
  xlab("Age") + 
  ylab("Amount of Bill in July") +
  coord_cartesian(xlim = c(21,60),ylim = c(0,700))+  
  geom_jitter(alpha=0.3, color="orange") + geom_smooth(stat='summary', fun.y=mean)

aug <- ggplot(aes(x=AGE,y=AUG2005/1000),data=ccdata) +
  xlab("Age") + 
  ylab("Amount of Bill in August") +
  coord_cartesian(xlim = c(21,60),ylim = c(0,700))+  
  geom_jitter(alpha=0.3, color="blue") + geom_smooth(stat='summary', fun.y=mean)

sep <- ggplot(aes(x=AGE,y=SEP2005/1000),data=ccdata) +
  xlab("Age") + 
  ylab("Amount of Bill in September") +
  coord_cartesian(xlim = c(21,60),ylim = c(0,700))+  
  geom_jitter(alpha=0.3, color="green") + geom_smooth(stat='summary', fun.y=mean)
  
grid.arrange(apr,may,jun,jul,aug,sep,ncol=3)

#Result: Distribution of expenditures on ages doesn’t show any differentiation between months and the average expenditure by age is almost same for each month.
```

### Correlations Between Limit Balance, Bill Amounts & Payments
```{r, comment=''}
M <- cor(subset(ccdata, select = c(LIMIT_BAL,BILL_AMT1,BILL_AMT2,BILL_AMT3,BILL_AMT4,BILL_AMT5,PAY_AMT1,PAY_AMT2,PAY_AMT3,PAY_AMT4,PAY_AMT5,PAY_AMT6)))
corrplot(M, method="number")

#Result: When we reflect the correlations between limit balances, bill amounts and payments amounts; it presents us that there’s a low correlation between the limit balances and payments and bill amounts. Howevet it can be seen that bill amounts has high correlation between each other as expected since the bills a reflecting the cumulative amounts.
```

### Personal Balance Limits Probabilities & Given Limits By Age
```{r, comment=''}
ggplot(aes(x=AGE,y=LIMIT_BAL/1000),data=subset(ccdata,!is.na(AGE.bucket)))+
  xlab("Age") + 
  ylab("Balance Limit (x1000 NT$)") +
  coord_cartesian(xlim = c(21,60),ylim = c(0,700))+
  scale_color_brewer(palette = "Pastel1")+
  geom_jitter(alpha=0.5, position = position_jitter(h=0), aes(color=AGE.bucket)) +
  geom_smooth(stat='summary', fun.y=mean) +
  geom_smooth(stat='summary', fun.y=quantile, fun.args = list(probs = 0.1), color = 'black', linetype=2) +
  geom_smooth(stat='summary', fun.y=quantile, fun.args = list(probs = 0.5), color = 'red', linetype=2) +
  geom_smooth(stat='summary', fun.y=quantile, fun.args = list(probs = 0.9), color = 'black', linetype=2)

#Result: When we plot the average balance limit per age bucket, we realized there are some limit levels given to clients such as 500,000 and 200,000 and these levels are belong to the probabilities of 95% and 50% respectively. So we can say that %95 of clients has balance limits equal or less than TL500k and %50 of clients has balance limits equal or less than TL200k.
# Additionally, when we investigate the blue line which represents the average limit balance of all clients, we think that we can interpret the spread between blue line and discrete red line as the bank prefers to give more balance limit than the average for the 50% probability.
# On the other hand, I can interpret the probabilities as the confidence intervals of exposures of bank to clients and I can say that bank implies that it can lose more than TL500k only by 5% if there is a client default occurs.
```

```{r, comment=''}
ggplot(aes(x=AGE,y=LIMIT_BAL/1000),data=subset(ccdata,!is.na(AGE.bucket)))+
  xlab("Age") + 
  ylab("Balance Limit (x1000 NT$)") +
  coord_cartesian(xlim = c(21,60),ylim = c(0,700))+
  scale_color_brewer(palette = "Pastel1")+
  geom_jitter(alpha=0.5, position = position_jitter(h=0), aes(color=AGE.bucket)) +
  geom_smooth(stat='summary', fun.y=mean) +
  geom_smooth(stat='summary', fun.y=quantile, fun.args = list(probs = 0.1), color = 'black', linetype=2) +
  geom_smooth(stat='summary', fun.y=quantile, fun.args = list(probs = 0.5), color = 'red', linetype=2) +
  geom_smooth(stat='summary', fun.y=quantile, fun.args = list(probs = 0.9), color = 'black', linetype=2) +
  geom_smooth(stat='summary', fun.y=quantile, fun.args = list(probs = 0.95), color = 'black', linetype=2) +
  facet_wrap(~default.payment.next.month)

#Result: Within the number of observations having a limit of 500,000 or less, it is not necessary to be more than 5% of the number of observations of the number of defaults when the default probability is related.
#Otherwise, 
#      a) A change in the default prediction model is required.
#      b) A change in the bank's limit allocation model is required.
# We can also say that when defining a limit, a limit of over 50% is defined, it is one of the reasons why people fall into default payment.
```